---
title: "data_cleaning"
author: "Xiaojie Mao"
date: "6/9/2019"
output: html_document
---


# Data downloading

## HMDA dataset 
We use the HMDA 2011-2012 data following the CFPB's technical report: https://files.consumerfinance.gov/f/201409_cfpb_report_proxy-methodology.pdf

The data can be downloaded from https://www.consumerfinance.gov/data-research/hmda/explore. We query the following observations and variables:

- where: (applicant_ethnicity=1 OR applicant_ethnicity=2) AND (applicant_race_1=1 OR applicant_race_1=2 OR applicant_race_1=3 OR applicant_race_1=4 OR applicant_race_1=5) AND (action_taken=1 OR action_taken=2 OR action_taken=3) AND (as_of_year=2011 OR as_of_year=2012)
- select: action_taken_name, applicant_ethnicity_name, applicant_income_000s, applicant_race_name_1, applicant_race_name_2, applicant_sex_name, as_of_year, census_tract_number, co_applicant_ethnicity_name, co_applicant_race_name_1, co_applicant_race_name_2, co_applicant_sex, county_name, loan_amount_000s, population, rate_spread, state_code
- See https://cfpb.github.io/api/hmda/fields.html for the meaning of the variables. 
- Copy paste the following link to the browser and it will start downloading the data (it takes a while to process the query and start downloading):
`https://api.consumerfinance.gov/data/hmda/slice/hmda_lar.csv?&$where=as_of_year+IN+(2012,2011)+AND+action_taken+IN+(1,2,3)+AND+applicant_race_1+IN+(1,2,3,4,5)+AND+applicant_ethnicity+IN+(1,2)&$select=action_taken_name,%20applicant_ethnicity_name,%20applicant_income_000s,%20applicant_race_name_1,%20applicant_race_name_2,%20applicant_sex_name,%20as_of_year,%20census_tract_number,%20co_applicant_ethnicity_name,%20co_applicant_race_name_1,%20co_applicant_race_name_2,%20co_applicant_sex,%20county_name,%20loan_amount_000s,%20population,%20rate_spread,%20state_code&$limit=0`

The dataset is around 3.5 G with name hmda_lar.csv. 

## Geolocation proxy
We use the census-tract level geolocation proxy (tract_over18_race_dec10.dta) that can be downloaded from https://github.com/cfpb/proxy-methodology 

# Data cleaning 
```{r cache = T}
# read the raw data 
library(tidyverse)
library(foreign)
library(totalcensus)
source("data_cleaning_helper.R")
hmda = read_csv("hmda_lar.csv")
hmda = as.tibble(hmda)
```

```{r}
# remove missing data 
hmda <- hmda %>%
  remove_obs(get_obs_state_missing) %>%  # remove observations whose state_code is missing
  remove_obs(get_obs_tract_missing) %>%  # remove observations whose tract number is missing
  remove_obs(get_obs_county_missing) %>%  # remove observations whose county is missing
  transform_tract_number()  # remove the . in census tract number: "1111.11" --> "111111"


```

```{r}
# Census Preprocesssing 
tract_over18 = read.dta("tract_over18_race_dec10.dta")
tract_over18 = select(tract_over18, -one_of(c("GEOID10_Tract", "Non_Hispanic_Total", "NH_Other_alone", "NH_Mult_Total", "NH_White_Other", "NH_Black_Other", "NH_AIAN_Other", "NH_Asian_HPI", "NH_API_Other", "NH_Asian_HPI_Other")))

tract_over18 = mutate(tract_over18, base_pop = Hispanic_Total + NH_White_alone + NH_Black_alone + NH_AIAN_alone + NH_API_alone) 
tract_over18 <- tract_over18 %>%
  mutate(Hispanic = Hispanic_Total/base_pop,
              White = NH_White_alone/base_pop,
              Black = NH_Black_alone/base_pop,
              AIAN = NH_AIAN_alone/base_pop,
              API = NH_API_alone/base_pop) %>% 
  select(-one_of(c("Hispanic_Total", "NH_White_alone", "NH_Black_alone", "NH_AIAN_alone", "NH_API_alone")))
tract_over18 = as.tibble(tract_over18)
tract_over18$State_FIPS10 = as.integer(tract_over18$State_FIPS10)
# write_csv(tract_over18, "BIG.csv")
```

```{r}
# combine the census tract proxy with the hmda dataset 
# tract_over18 = read_csv("BIG.csv")

# get the county code for the hmda dataset (originally use county names)
fips = search_fips(view = F)
fips_county <- fips %>% 
  select(one_of(c("STATE", "COUNTY", "NAME")))
fips_county$STATE = as.integer(fips_county$STATE)

temp = hmda %>% 
  left_join(fips_county, by = c("state_code"="STATE", "county_name"="NAME")) %>% 
  filter(!is.na(COUNTY))

# combine the census tract proxy and the hmda dataset 
temp <- temp %>% 
  left_join(tract_over18, by = c("state_code"="State_FIPS10", "COUNTY"="County_FIPS10", "census_tract_number"="Tract_FIPS10")) %>% 
  filter(!is.na(White))

temp = temp %>% select(applicant_ethnicity_name, applicant_race_name_1, applicant_race_name_2, applicant_income_000s, census_tract_number, COUNTY, state_code, 
                Hispanic, White, Black, AIAN, API, outcome)

# write_csv(temp, "hmda_census_combined.csv")
```




